---
title: "Pareto Plots 2"
author: "Bruce Hartman"
output: html_notebook
---

## Pareto Plots
Pareto plots are a good way to understand what's important in creating unicorns or increasing their valuation. We use the summarized data frames for Count and TotVal.
```{r}
glimpse(x)
x
```

```{r}
glimpse(y)
y
```

## Pareto Functions for Plotting
We use the <code>qcc</code> package to create Pareto chart data from the count data. Then we use <code>ggQC</code> package to create Pareto graphs of Total Valuation by year. We define two functions in a separate file.

- <code>pareto_data_structure</code> creates a Pareto data structure 
- <code>pareto_plot_object</code> creates a Pareto plot object from the structure.

```{r}
source("paretoplotbuilder.R")
```

## Pareto Plots of Count
Pareto Plots of Count for 2015 through 2021.
```{r}
for (yr in 2015:2021) {
  uu = pareto_data_structure(y, yr, which="Count") %>% remove_rownames
  uu%>% filter(CumPct<80)
  p=pareto_plot_object(uu, yr, which="Count", main="Pareto Plot")
  print(p)
  # ggsave(paste0("Innovation/Pareto Plot Count Year ",yr,".jpg"), plot=p, device="jpeg")
}
```

## Save Pareto Plots of TotVal
```{r}
for (yr in 2015:2021) {
  uu = pareto_data_structure(x, yr, which="TotVal") %>% remove_rownames
  uu%>% filter(CumPct<80)
  p=pareto_plot_object(uu, yr, which="TotVal", main="Pareto Plot")
  print(p)
  # ggsave(paste0("Innovation/Pareto Plot TotVal Year ",yr,".jpg"), plot=p, device="jpeg")
}
```



## Total Valuation by Class and Year
A barplot will be useful.
```{r}
pv=ggplot(x %>% filter(between(Year,2016,2021) )) + 
  geom_col(aes(Class, TotVal, fill=Class)) +
  scale_fill_manual(values=c("blue","yellow","red")) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
  labs(title="Total Valuation by Class and Year") +
  facet_wrap(vars(Year), ncol=2)
pv
# ggsave("Innovation/Valuation by Class and Year.jpg", plot=pv, device="jpeg", height=7 )
```

## Count by Class and Year
A barplot will be useful.
```{r}
pv=ggplot(y %>% filter(between(Year,2016,2021) )) + 
  geom_col(aes(Class, Count, fill=Class)) +
  scale_fill_manual(values=c("blue","yellow","red")) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
  labs(title="Count by Class and Year") +
  facet_wrap(vars(Year), ncol=2)
pv
# ggsave("Innovation/Count by Class and Year.jpg", plot=pv, device="jpeg", height=7 )
```


## Valuations by Year and Country
```{r}
for (yr in 2016:2021){
  p = ggplot(cbi_more %>% group_by(Country, Class, sum(Valuation)) %>% filter(Year==yr)) + 
    geom_col(aes(Country,sum(Valuation), fill=Class), alpha=.6, position="stack") +
    scale_fill_manual(values=c("blue","yellow","red")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=.5)) +
    labs(title=paste0("Total Valuation by Country Year ",yr))
  print(p)
  # ggsave(paste0("Innovation/TotVal by Country Year ",yr,".jpg"), plot=p, device = "jpeg", height=7, width=10)
}
```

Next we leave out the USA because it is so much bigger in later years.
```{r}
for (yr in 2016:2021){
  p = ggplot(cbi_more %>% group_by(Country, Class, sum(Valuation)) %>% filter(Year==yr, Country!="United States")) + 
    geom_col(aes(Country,sum(Valuation), fill=Class), alpha=.6, position="stack") +
    scale_fill_manual(values=c("blue","yellow","red")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=.5)) +
    labs(title=paste0("Total Valuation by Country x-USA Year ",yr))
  print(p)
  # ggsave(paste0("Innovation/TotVal by Country x-USA, Year ",yr,".jpg"), plot=p, device = "jpeg", height=7, width=10)
}
```

Arrange plots in a grid and plot.This is not too effective.
```{r}
plo = list()
for (yr in 2016:2021){
  p = ggplot(cbi_more %>% group_by(Country, Class, sum(Valuation)) %>% filter(Year==yr, Country!="United States")) + 
    geom_col(aes(Country,sum(Valuation), fill=Class), alpha=.6, position="stack") +
    scale_fill_manual(values=c("blue","yellow","red")) +
    theme(axis.text.x = element_text(angle = 90, vjust = .9, hjust=-.2)) +
    labs(#subtitle=paste0(#"Total Valuation by Country x-USA Year ",
        # yr),
         x="", y=paste0("Valuation ",yr))
  plo[[as.character(yr)]]=p
  # print(p)
  # ggsave(paste0("Innovation/TotVal by Country x-USA, Year ",yr,".jpg"), plot=p, device = "jpeg", height=7, width=10)
}
gr = lemon::grid_arrange_shared_legend(# plo$`2016`, plo$`2017`, 
                                  plo$`2018`,
                                  plo$`2019`, plo$`2020`, plo$`2021`, ncol=2, nrow=2, position="right")
gr
# ggsave("Innovation/Valuation by Country xUS 2018-2021.jpg", device = "jpeg", width=7, height=9, units="in")
```
